My title

### 2024-11-04 paper/paper.qmd
Added Appendix

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Abdullah Motasim
  - Elizabeth Luong
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election](https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
table-of-contents: true
toc-depth: 2
bibliography: references.bib
editor: 
  markdown: 
    wrap: sentence
---

@@

Estimand paragraph: The primary estimand in this study is the probability of a Democratic or Republican victory in the popular vote, based on current polling trends and demographic predictors.

Results paragraph: State results of who won in which state (the 5 in exploratory) and also talk about the final model and who it predicted to win in each state on the day of the election than also who won overall.

Why it matters paragraph:

Telegraphing paragraph: The remainder of this paper is structured as follows.
@sec-data....

@@

Note this is a general overview of the measurement method utilized by Emerson, a more thorough breakdown of measurement can be found in \[Appendix A\].

## Outcome variables

@@

State: Including the state where each poll was conducted allows us to account for regional voting patterns, which are essential in a federal system like the U.S.
Electoral College.
State-level data provides insights into geographic trends and helps us model potential outcomes across electoral regions.

@@

# Model

@@

Background details and diagnostics are included in [Appendix -@sec-model-details].

@@

The summary statistics below @SummaryStatisticsComparison outline key variables in our analysis, focusing on Kamala Harris and Donald Trump.
These tables display the distribution of support percentages and sample sizes, with metrics including mean, median, standard deviation, minimum, and maximum.

```{r}
#| echo: false
#| message: false
#| warning: false
#| label: SummaryStatisticsComparison
#| tbl-cap: Directly compares support percentages and sample sizes for Harris and Trump, showing central tendencies and variability in polling data for each candidate.

@@

# Results

@@

@@

The distribution of predicted support for Kamala Harris is displayed in Figure 4.
The model estimates her average support at approximately 47.32%, with the majority of values clustering around this mean.
This relatively tight distribution indicates consistent support for Harris across polls, with limited variability.
The standard deviation of her predicted support suggests that her polling results remain stable, even in states with more diverse polling outcomes.

### Donald Trump

Figure 5 illustrates the predicted support for Donald Trump, centered around a mean of 45.56%.
Compared to Harris, Trump’s predicted support distribution shows slightly greater variability, indicating a broader spread in his support across polls.
The higher standard deviation in Trump’s support highlights increased variability in polling outcomes, suggesting a wider range of voter responses depending on the region and poll characteristics.

@@

The combined density plot in Figure 6 provides a direct comparison of Harris and Trump’s predicted support distributions.
Harris’s distribution (in blue) peaks at a higher level than Trump’s (in red), indicating her slight edge in predicted support.
Both distributions overlap significantly, particularly between 45% and 50%, underscoring the close nature of their polling outcomes.
This overlap suggests that while Harris holds a predicted lead, the margins are slim, indicating a competitive scenario in many states.

@@

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion - Predictive Modeling of 2024 Election Outcomes

@@

# Pollster Methodology Overview and Evaluation

@@

[**U.S. Presidential Election Forecasting Survey**](https://docs.google.com/forms/d/e/1FAIpQLSdlSVq77BK3tKqsrsdLZicLx8ie0Jei1PFX3HzOGmpKxolD8w/viewform?usp=sf_link)

**Survey Copy**

**U.S. Presidential Election Forecasting Survey**

@@

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check.
This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior.
This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
# 
# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot.
It shows...
This suggests...

@fig-stanareyouokay-2 is a Rhat plot.
It shows...
This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Abdullah Motasim
  - Elizabeth Luong
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election](https://github.com/abdullah-motasim/Forecasting-2024-US-Presidential-Election)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
table-of-contents: true
toc-depth: 2
bibliography: references.bib
---

@@

The primary estimand in this study is the probability of a Democratic or Republican victory in the popular vote, based on current polling trends and demographic predictors.

Utilizing Bayesian Generalized Linear Models (GLM) and Monte Carlo simulations to estimate the predicted support percentages for Kamala Harris and Donald Trump. Harris has a higher support percentage over Trump with a win probability of 63.2%, while Trump has a 36.8% win probability. The model suggest that although there is close competition between both candidates, Harris has an advantage based on the polling data.

Running the simulations for this data showcases the importance of state-level effects, support for each candidate varies significantly by region. The models indicate that Harris has strong support in traditionally Democratic states, whereas Trump maintains a lead in Republican-leading regions. Support for each candidate indicates that regional factors are pivotal in determining the final outcome for American elections.

The remainder of this paper is structured as follows.
@sec-data discusses the data types included in the raw data, the cleaning process for the data, and the reason for selecting the data set we did. @sec-model discusses model specification and justification for Bayesian GLMs and Monte Carlo simulation. @sec-results analyzes the trends and correlations between different variables utilizing tabular and graphical means. @sec-discussion discusses the results of @sec-results going into detail on what the simulation results can tell us about the upcoming election, as well as assumptions and limitations in data.

@@

Note this is a general overview of the measurement method utilized by Emerson, a more thorough breakdown of measurement can be found in @sec-pollster-methdology.

## Outcome variable{#sec-outcome-variable}

@@

State: Including the state where each poll was conducted allows us to account for regional voting patterns, which are essential in a federal system like the U.S.

@@

# Model {#sec-model}

@@

@@

The summary statistics below @tbl-summary-statistics-comparison outline key variables in our analysis, focusing on Kamala Harris and Donald Trump.
These tables display the distribution of support percentages and sample sizes, with metrics including mean, median, standard deviation, minimum, and maximum.

```{r}
#| echo: false
#| message: false
#| label: tbl-summary-statistics-comparison
#| tbl-cap: Directly compares support percentages and sample sizes for Harris and Trump, showing central tendencies and variability in polling data for each candidate.

@@

# Results {#sec-results}

@@

```{r}
#| echo: false
#| warning: false
#| message: false
#| label: fig-harris-support
#| tbl-cap: This histogram shows the distribution of predicted support percentages for Kamala Harris based on model simulations.

# Plot distribution of predicted support for Harris
ggplot(data.frame(harris_pred_means), aes(x = harris_pred_means)) +
  geom_histogram(bins = 30, fill = "blue", color = "black", alpha = 0.7) +
  labs(
    title = "Distribution of Predicted Support \nfor Kamala Harris",
    x = "Predicted Support Percentage",
    y = "Frequency",
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.caption = element_text(size = 10, hjust = 0.5, margin = margin(t = 10))
  )

@@

The distribution of predicted support for Kamala Harris is displayed in @fig-harris-support.
The model estimates her average support at approximately 47.32%, with the majority of values clustering around this mean.
This relatively tight distribution indicates consistent support for Harris across polls, with limited variability.
The standard deviation of her predicted support suggests that her polling results remain stable, even in states with more diverse polling outcomes.

@@

The combined density plot was @fig-avg-support previously shown @sec-outcome-variable provides a direct comparison of Harris and Trump’s predicted support distributions.
Harris’s distribution (in blue) peaks at a higher level than Trump’s (in red), indicating her slight edge in predicted support.
Both distributions overlap significantly, particularly between 45% and 50%, underscoring the close nature of their polling outcomes.
This overlap suggests that while Harris holds a predicted lead, the margins are slim, indicating a competitive scenario in many states.

@@

# Discussion - Predictive Modeling of 2024 Election Outcomes {#sec-discussion}

@@

# Pollster Methodology Overview and Evaluation {#sec-pollster-methdology}

@@

**Survey Link:** [U.S. Presidential Election Forecasting Survey](https://docs.google.com/forms/d/e/1FAIpQLSdlSVq77BK3tKqsrsdLZicLx8ie0Jei1PFX3HzOGmpKxolD8w/viewform?usp=sf_link)

**Survey Copy: **

@@
